Learning From Massive Noisy Labeled Data for Image Classification

### 2015-04-10 ext.tex
Add extended abstract

%!TEX program = xelatex
\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{algorithm2e}
\usepackage{authblk}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

\usepackage{fontspec}

\cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{407} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

\def\vec{\mathbf}
\def\mat{\mathbf}

\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}

% Pages are numbered in submission mode, and unnumbered in camera-ready
\ifcvprfinal\pagestyle{empty}\fi
\begin{document}

\RestyleAlgo{boxruled}

%%%%%%%%% TITLE
\title{Learning From Massive Noisy Labeled Data for Image Classification}

\author[1]{Tong Xiao}
\author[2]{Tian Xia}
\author[2]{Yi Yang}
\author[2]{Chang Huang}
\author[1]{Xiaogang Wang}

\affil[1]{The Chinese University of Hong Kong}
\affil[2]{Baidu Research}

\renewcommand\Authands{, and }

% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both

\maketitle
%\thispagestyle{empty}

%%%%%%%%% ABSTRACT
\begin{abstract}
Large-scale supervised datasets are crucial to train convolutional neural networks (CNNs) for various computer vision problems. However, obtaining a massive amount of well labeled data is usually very expensive and time consuming. In this paper, we introduce a general framework to train CNNs with only a limited number of clean labels and millions of easily obtained noisy labels. We describe label noises with a probabilistic graphical model and further integrate it into an end-to-end deep learning system. To demonstrate the effectiveness of our approach, we collect a large-scale clothing classification dataset with both noisy and clean labels. Experiments on this dataset show that our model can better detect and correct the wrong labels, which benefits the training of underlying CNNs.
\end{abstract}

%%%%%%%%% BODY TEXT
\section{Introduction} % (fold)
\label{sec:introduction}

Deep learning with a massive amount of supervised training data has recently shown very impressive improvement on multiple image recognition challenges including image classification~\cite{krizhevsky2012imagenet}, attribute learning~\cite{zhang2014panda}, and scene classification~\cite{farabet2013learning}. While state-of-the-art results have been continuously reported~\cite{zeiler2013visualizing,simonyan2014very,szegedy2014going}, all these methods require reliable annotations from millions of images~\cite{deng2009imagenet} which are often expensive and time-consuming to obtain~\cite{deng2009imagenet}, preventing deep models from being quickly trained on new image recognition problems. Thus it is necessary to develop new efficient labeling and training frameworks for deep learning.

One possible solution is to automatically collect large amount of annotations from the Internet web images~\cite{gong2014multiview} (i.e. extracting tags from the surrounding texts or keywords from search engines) and directly use them as ground truth to train deep models. Unfortunately, these labels are extremely unreliable due to various types of noise (i.e. labeling mistakes from annotators or computing errors from extraction algorithms). Many works have shown that these noisy labels could adversely impact the classification accuracy of the induced classifiers~\cite{zhu2004class,nettleton2010study,pechenizkiy2006class}. Various label noise-robust algorithms are developed but experiments show that performances of classifiers inferred by robust algorithms are still affected by label noise~\cite{bartlett2006convexity,teng2001comparison}. Other data cleansing algorithms are proposed~\cite{barandela2000decontamination,brodley2011identifying,miranda2009use}, but these approaches are difficult in distinguishing informative hard examples from harmful mislabeled ones.

\begin{figure}[t]
\begin{center}
\includegraphics[width=0.95\linewidth]{figure/overview.pdf}
\end{center}
\caption{Overview of our approach. Labels of web images often suffer from different types of noise. A label noise model is proposed to detect and correct the wrong labels. The corrected labels are used to train underlying CNNs.}
\label{fig:overview}
\end{figure}

Although annotating all the data is costly, it is often easy to obtain a small amount of clean labels. Based on the observation of transferability of deep neural networks, people initialize parameters with a model pretrained on a larger yet related dataset~\cite{krizhevsky2012imagenet}, and then finetune on the smaller dataset of specific tasks~\cite{oquab2014learning,azizpour2014generic,donahue2013decaf}. Such methods may better avoid overfitting and utilize the relationships between the two datasets. However, we find that training a CNN from scratch with limited clean labels and massive noisy labels is better than finetuning it only on clean labels. Other approaches address the problem as semi-supervised learning where noisy labels are discarded~\cite{zhu2002learning}. These algorithms usually suffer from model complexity thus cannot be applied on large-scale datasets. Therefore, it is inevitable to develop a better way of using huge amount of noisy labeled data.

@@

To cope with this challenge, we extend CNNs with a novel probabilistic model, which infers the true labels and uses them to supervise the training of the network. Our work is inspired by~\cite{sukhbaatar2014learning}, which modified a CNN by inserting a linear layer on top of the softmax layer to map clean labels to noisy labels. However,~\cite{sukhbaatar2014learning} assumed noisy labels were conditionally independent with input images given clean labels. By examining our collected dataset, we find that this assumption is too strong to fit real-world data well. For example, in Figure~\ref{fig:noise_cluster}, all the images should belong to ``Hoodie''. The top five are correct while the bottom five are either mislabeled as ``Windbreaker'' or ``Jacket''. This is because different sellers have their own bias on different categories, thus they may provide wrong keywords for similar clothes. We can capture these visual patterns to estimate how likely an image is mislabeled. Based on these observations, we further introduce two types of label noise:
\begin{itemize}
\item \textbf{Confusing noise} makes the noisy label reasonably wrong. It usually occurs when the image content is confusing (\eg, the samples with ``?'' in Figure~\ref{fig:overview}).
\item \textbf{Pure random noise} makes the noisy label totally wrong. It is often caused by either the mismatch between an image and its surrounding text, or false conversion from the text to label (\eg, the samples with ``$\times$'' in Figure~\ref{fig:overview}).
\end{itemize}

The proposed probabilistic model builds the relations among images, noisy labels, ground truth labels, and noise types, where the latter two are treated as latent variables. We use the Expectation-Maximization (EM) algorithm to solve the problem and integrate it into the training process of CNNs. Experiments on the collected dataset show that our model can better detect and correct the wrong labels.

Our contribution comes from three aspects. First, we study the cause of noisy labels in real-world data and describe it with a novel probabilistic model. Second, we integrate the model into a deep learning framework to help detect and correct wrong labels. We study different training strategies to make the CNNs learn from better supervisions. Finally, we collect a large-scale clothing dataset with both noisy and clean labels, which will be released for academic use.

\begin{figure}[t]
\begin{center}
\includegraphics[width=1.0\linewidth]{figure/noise_cluster.pdf}
\end{center}
\caption{Mislabeled images often share similar visual patterns.}
\label{fig:noise_cluster}
\end{figure}
% section introduction (end)

\section{Related Work} % (fold)
\label{sec:related_work}
For most of the related works including the effect of label noises, taxonomy of label noises, robust algorithms and noise cleaning algorithms for learning with noisy data, we refer to~\cite{frenay2014classification} for a comprehensive review.

{\bf Direct learning with noisy labels:} Many works have shown that label noise can adversely impact the classification accuracy of induced classifiers~\cite{zhu2004class}. To better handle label noise, some approaches rely on training classifiers with label noise-robust algorithms~\cite{beigman2009learning,manwani2013noise}. However, Bartlett \etal~\cite{bartlett2006convexity} proved that most of the loss functions are not completely robust to label noise. Experiments in~\cite{teng2001comparison} showed that the classifiers inferred by label noise-robust algorithms are still affected by label noise. These methods seem to be adequate only when label noise can be safely managed by overfitting avoidance~\cite{frenay2014classification}. On the other hand, some label noise cleansing methods were proposed to remove or correct mislabeled instances~\cite{barandela2000decontamination,brodley2011identifying,miranda2009use}, but these approaches were difficult in distinguishing informative hard examples from harmful mislabeled ones. Thus they might remove too many instances and the overcleansing could reduce the performances of classifiers~\cite{matic1992computer}.

{\bf Semi-supervised learning:} Apart from direct learning with label noise, some semi-supervised learning algorithms were developed to utilize weakly labeled or even unlabeled data. The Label Propagation method~\cite{zhu2002learning} explicitly used ground truths of well labeled data to classify unlabeled samples. However, it suffered from computing pairwise distance, which has quadratic complexity with the number of samples thus cannot be applied on large-scale datasets. Weston \etal~\cite{weston2012deep} proposed to embed a pairwise loss in the middle layer of a deep neural network, which benefits the learning of discriminative features. But they needed extra information about whether a pair of unlabeled images belong to the same class, which cannot be obtained in our problem.

{\bf Transfer learning:} The success of CNNs lies in their capability of learning rich and hierarchical image features. However, the model parameters cannot be properly learned when training data is not enough. Researchers proposed to conquer this problem by first initializing CNN parameters with a model pretrained on a larger yet related dataset, and then finetuning it on the smaller dataset of specific task~\cite{krizhevsky2012imagenet, oquab2014learning, azizpour2014generic,donahue2013decaf}. Nevertheless, this transfer learning scheme could be suboptimal when the two tasks are just loosely related. In our case of clothing classification, we find that training a CNN from scratch with limited clean labels and massive noisy labels is better than finetuning it only on the clean labels.

{\bf Noise modeling with deep learning:} Various methods have been proposed to handle label noise in different problem settings, but there are very few works about deep learning from noisy labels~\cite{mnih2012learning, larsen1998design, sukhbaatar2014learning}. Mnih and Hinton~\cite{mnih2012learning} built a simple noise model for aerial images but only considered binary classification. Larsen \etal~\cite{larsen1998design} assumed label noises are independent from true class labels which is a simple and special case. Sukhbaatar \etal~\cite{sukhbaatar2014learning} generalized from them by considering multi-class classification and modeling class dependent noise, but they assumed the noise was conditionally independent with the image content, ignoring the hardness of labeling images of different confusing levels. Our work can be viewed as a generalization of~\cite{sukhbaatar2014learning,natarajan2013learning} and our model is flexible enough to not only class dependent noise but also input dependent noise.
% section related_work (end)

\section{Label Noise Model} % (fold)
\label{sec:label_noise_model}
We target on learning a classifier from a set of images with noisy labels. To be specific, we have a noisy labeled dataset $\mathcal{D}_{\eta}=\left\{\left(\vec{x}^{(1)}, \tilde{y}^{(1)}\right), \dots, \left(\vec{x}^{(N)}, \tilde{y}^{(N)}\right)\right\}$ with $n$-th image $\vec{x}^{(n)}$ and its corresponding noisy label $\tilde{y}^{(n)} \in \{1,\dots,L\}$, where $L$ is the number of classes. We describe how the noisy label is generated by using a probabilistic graphical model shown in Figure~\ref{fig:noise_pgm}.

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.5\linewidth]{figure/pgm.pdf}
\end{center}
\caption{Probabilistic graphical model of label noise}
\label{fig:noise_pgm}
\end{figure}

Despite the observed image $\vec{x}$ and the noisy label $\tilde{\vec{y}}$, we exploit two discrete latent variables --- $\vec{y}$ and $\vec{z}$ --- to represent the true label and the label noise type, respectively. Both $\tilde{\vec{y}}$ and $\vec{y}$ are $L$-dimensional binary random variables in 1-of-$L$ fashion, \ie, only one element is equal to 1 while others are all 0.

The label noise type $\vec{z}$ is an 1-of-3 binary random variable. It is associated with three different semantic meanings:
\begin{enumerate}
    \item The label is noise free, \ie, $\tilde{\vec{y}}$ should be equal to $\vec{y}$
    \item The label suffers from a pure random noise, \ie, $\tilde{\vec{y}}$ can take any possible value other than $\vec{y}$
    \item The label suffers from a confusing noise, \ie, $\tilde{\vec{y}}$ can take several values that are confusing with $\vec{y}$
\end{enumerate}

Following this assignment rule, we define the conditional probability of the noisy label as
\begin{equation} \label{eq:likelihood}
    p(\tilde{\vec{y}}|\vec{y},\vec{z}) = \begin{cases}
        \tilde{\vec{y}}^{T}\mat{I}\vec{y} & \text{if } \vec{z}_1 = 1 \\
        \frac{1}{L-1}\tilde{\vec{y}}^{T}(\mat{U} - \mat{I})\vec{y} & \text{if } \vec{z}_2 = 1 \\
        \tilde{\vec{y}}^{T}\mat{C}\vec{y} & \text{if } \vec{z}_3 = 1
    \end{cases}
\end{equation}
where $\mat{I}$ is the identity matrix, $\mat{U}$ is the unit matrix (all the elements are ones), $\mat{C}$ is a sparse stochastic matrix with $\mathrm{tr}(\mat{C})=0$ and $\mat{C}_{ij}$ denotes the confusion probability between classes $i$ and $j$. Then we can derive from Figure~\ref{fig:noise_pgm} the joint distribution of $\tilde{\vec{y}}, \vec{y}$ and $\vec{z}$ conditioning on $\vec{x}$:
\begin{equation} \label{eq:joint_distribution}
  p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}) = p(\tilde{\vec{y}} | \vec{y}, \vec{z}) p(\vec{y} | \vec{x}) p(\vec{z} | \vec{x})
\end{equation}

While the class label probability distribution $p(\vec{y} | \vec{x})$ is comprehensible, the semantic meaning of $p(\vec{z} | \vec{x})$ needs extra clarification: it represents how confusing the image content is. Specific to our clothing classification problem, $p(\vec{z} | \vec{x})$ can be affected by different factors, including background clutter, image resolution, the style and material of the clothes. Some examples are shown in Figure~\ref{fig:compare_noise_level}.

\begin{figure}[t]
\begin{center}
\includegraphics[width=1.0\linewidth]{figure/noise_level.pdf}
\end{center}
\caption{Predicting noise types of four different ``T-shirt'' images. The top two can be recognized with little ambiguity, while the bottom two are easily confusing with the class ``Chiffon''. Image content can affect the possibility of it to be mislabeled.}
\label{fig:compare_noise_level}
\end{figure}

To illustrate the relations between noisy and true labels, we derive their conditional probability from Eq~\ref{eq:joint_distribution}
\begin{equation}
    p(\tilde{\vec{y}} | \vec{y}, \vec{x}) = \sum_{\vec{z}} p(\tilde{\vec{y}}, \vec{z} | \vec{y}, \vec{x}) = \sum_{\vec{z}} p(\tilde{\vec{y}} | \vec{y}, \vec{z}) p(\vec{z} | \vec{x})
\end{equation}
which can be interpreted as a mixture model. Given an input image $\vec{x}$, the conditional probability $p(\vec{z} | \vec{x})$ can be seen as the prior of each mixture component. This makes a key difference between our work and~\cite{sukhbaatar2014learning}, where they assume $\tilde{\vec{y}}$ is conditionally independent with $\vec{x}$ if $\vec{y}$ is given. All the images share a same noise model in~\cite{sukhbaatar2014learning}, while in our approach each data sample has its own.

\subsection{Learning the Parameters} % (fold)
\label{sub:learning_the_parameters}
We exploit two CNNs to model $p(\vec{y} | \vec{x})$ and $p(\vec{z} | \vec{x})$ separately. Denote the parameter set of each CNN by $\theta_1$ and $\theta_2$. Our goal is to find the optimal $\theta = \theta_1 \cup \theta_2$ that maximize the incomplete log-likelihood $\log p(\tilde{\vec{y}} | \vec{x}; \theta)$. The EM algorithm is used to iteratively solve this problem.

For any probability distribution $q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x})$, we can derive a lower bound of the incomplete log-likelihood,
\begin{equation}
\begin{aligned}
    \log p(\tilde{\vec{y}} | \vec{x}; \theta) &= \log \sum_{\vec{y},\vec{z}} p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta) \\
    &\geq \sum_{\vec{y},\vec{z}} q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}) \log \frac{p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta)}{q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x})}
\end{aligned}
\end{equation}
\textbf{E-Step} The difference between $\log p(\tilde{\vec{y}} | \vec{x}; \theta)$ and its lower bound is the Kullback-Leibler divergence $\mathrm{KL}\left(q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}) || p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta)\right)$, which is equal to zero if and only if $q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}) = p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta)$. Therefore, in each iteration $t$, we first compute the posterior of latent variables using the current parameters $\theta^{(t)}$:
\begin{multline} \label{eq:posterior}
    p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) = \frac{p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta^{(t)})}{p(\tilde{\vec{y}} | \vec{x}; \theta^{(t)})} \\
    = \frac{p(\tilde{\vec{y}}|\vec{y},\vec{z};\theta^{(t)}) p(\vec{y}|\vec{x};\theta^{(t)}) p(\vec{z}|\vec{x};\theta^{(t)})}{\sum_{\vec{y}',\vec{z}'} p(\tilde{\vec{y}}|\vec{y}',\vec{z}';\theta^{(t)}) p(\vec{y}'|\vec{x};\theta^{(t)}) p(\vec{z}'|\vec{x};\theta^{(t)})}
\end{multline}
Then the expected complete log-likelihood can be written as
\begin{equation}
    Q(\theta; \theta^{(t)}) = \sum_{\vec{y},\vec{z}} p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) \log p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta)
\end{equation}
\textbf{M-Step} We exploit two CNNs to model the probability $p(\vec{y}|\vec{x};\theta_1)$ and $p(\vec{z}|\vec{x};\theta_2)$, respectively. Thus the gradient of $Q$ \wrt $\theta$ can be decoupled into two parts:
\begin{multline} \label{eq:gradient}
    \frac{\partial Q}{\partial \theta} = \sum_{\vec{y},\vec{z}} p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) \frac{\partial}{\partial \theta} \log p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta) \\
    = \sum_{\vec{y}} p(\vec{y} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) \frac{\partial}{\partial \theta_1} \log p(\vec{y}|\vec{x};\theta_1) + \\
      \sum_{\vec{z}} p(\vec{z} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) \frac{\partial}{\partial \theta_2} \log p(\vec{z}|\vec{x};\theta_2)
\end{multline}

The M-Step above is equivalent to minimizing the cross entropy between the estimated ground truth distribution and the prediction of the classifier.
% subsection learning_the_parameters (end)

\subsection{Estimating Matrix $\mat{C}$} % (fold)
\label{sub:confusion_estimation}
Notice that we do not set parameters to the conditional probability $p(\tilde{\vec{y}}|\vec{y},\vec{z})$ in Eq~\eqref{eq:likelihood} and keep it unchanged during the learning process. Because without other regularizations, learning all the three parts in Eq~\eqref{eq:joint_distribution} could lead to trivial solutions. For example, the network will always predict $\vec{y}_1 = 1$, $\vec{z}_3 = 1$, and the matrix $\mat{C}$ is learned to make $\mat{C}_{1i} = 1$ for all $i$. To avoid such degeneration, we estimate $\mat{C}$ on a relatively small dataset $\mathcal{D}_c = \{(\vec{x},\vec{y},\tilde{\vec{y}})\}_N$, where we have $N$ images with both clean and noisy labels. As prior information about $\vec{z}$ is not available, we solve the following optimization problem:
\begin{equation} \label{eq:opt_all}
    \max_{\mat{C},\vec{z}^{(1)},\cdots,\vec{z}^{(N)}} \sum_{i=1}^N \log p(\tilde{\vec{y}}^{(i)}|\vec{y}^{(i)},\vec{z}^{(i)})
\end{equation}
Obviously, sample $i$ contributes nothing to the optimal $\mat{C}^*$ if $\vec{y}^{(i)}$ and $\tilde{\vec{y}}^{(i)}$ are equal. So that we discard those samples and reinterpret the problem in another form by exploiting Eq~\ref{eq:likelihood}:
\begin{equation}
\begin{aligned}
\max_{\mat{C},\vec{t}} \quad & E=\sum_{i=1}^{N'} \log \alpha^{\vec{t}_i} + \log (\tilde{\vec{y}}^{(i)T} \mat{C} \vec{y}^{(i)})^{1-\vec{t}_i}\\
\text{subject to} \quad & \mat{C} \text{ is a stochastic matrix of size } L\times L \\
                    & \vec{t} \in \{0,1\}^{N'}
\end{aligned}
\end{equation}
where $\alpha=\frac{1}{L-1}$ and $N'$ is the number of remaining samples. The semantic meaning of the above formulation is that we need to assign each $(\vec{y},\tilde{\vec{y}})$ pair the optimal noise type, while finding the optimal $\mat{C}$ simultaneously.

Next, we will show that the problem can be solved by a simple yet efficient algorithm in $O(N'+L^2)$ time complexity. Denote the optimal solution by $\mat{C}^*$ and $\vec{t}^*$
\begin{theorem}
$\mat{C}^*_{ij} \neq 0 \Rightarrow \mat{C}^*_{ij} > \alpha, \forall i,j \in \{1,\dots,L\}$
\end{theorem}
\begin{proof}
Suppose there exists some $i,j$ such that $0 < \mat{C}^*_{ij} \leq \alpha$. Then we conduct the following operations. First, we set $\mat{C}^*_{ij}=0$ while adding its original value to other elements in column $j$. Second, for all the samples $n$ where $\tilde{\vec{y}}^{(n)}_i=1$ and $\vec{y}^{(n)}_j=1$, we set $\vec{t}_n$ to 1. The resulting $E$ will get increased, which leads to a contradiction.
\end{proof}
\begin{theorem}
$(\tilde{\vec{y}}^{(i)},\vec{y}^{(i)})=(\tilde{\vec{y}}^{(j)},\vec{y}^{(j)}) \Rightarrow \vec{t}^*_i=\vec{t}^*_j, \forall i,j \in \{1,\dots,N'\}$
\end{theorem}
\begin{proof}
Suppose $\tilde{\vec{y}}^{(i)}_k=\tilde{\vec{y}}^{(j)}_k=1$ and $\vec{y}^{(i)}_l=\vec{y}^{(j)}_l=1$ but $\vec{t}^*_i \neq \vec{t}^*_j$. From Theorem 1 we know that elements in $\mat{C}^*$ is either 0 or greater than $\alpha$. If $\mat{C}^*_{kl}=0$, we can set $\vec{t}^*_i=\vec{t}^*_j=1$, otherwise we can set $\vec{t}^*_i=\vec{t}^*_j=0$. In either case the resulting $E$ will get increased, which leads to a contradiction.
\end{proof}
\begin{theorem}
$\tilde{\vec{y}}^{(i)T} \mat{C}^* \vec{y}^{(i)} > \alpha \Leftrightarrow \vec{t}^*_i=0$ and $\tilde{\vec{y}}^{(i)T} \mat{C}^* \vec{y}^{(i)} = 0 \Leftrightarrow \vec{t}^*_i=1, \forall i \in \{1,\dots,N'\}$
\end{theorem}
\begin{proof}
The first part is straightforward. For the second part, $\vec{t}^*_i=1$ implies $\tilde{\vec{y}}^{(i)T} \mat{C}^* \vec{y}^{(i)} \leq \alpha$. By using Theorem 1 we know that $\tilde{\vec{y}}^{(i)T} \mat{C}^* \vec{y}^{(i)} = 0$.
\end{proof}

Notice that if the true label is class $i$ while the noisy label is class $j$, then it can only affect the value of $\mat{C}_{ij}$. Thus each column of $\mat{C}$ can be optimized separately. Theorem 1 further shows that samples with same pair of $(\tilde{\vec{y}},\vec{y})$ share a same noise type. Then what really matters is the frequencies of all the $L\times L$ pairs of $(\tilde{\vec{y}},\vec{y})$. Considering a particular column $\vec{c}$, suppose there are $M$ samples affecting this column. We can count the frequencies of noisy label class 1 to $L$ as $m_1,\dots,m_L$ and might as well assume $m_1\geq m_2\geq \cdots \geq m_L$. The problem is then converted to
\begin{equation}
\begin{aligned}
\max_{\vec{c},\vec{t}} \quad & E=\sum_{k=1}^{L} m_k \left(\log \alpha^{\vec{t}_k} + \log \mat{c}_k^{1-\vec{t}_k}\right),\\
\text{subject to} \quad & \vec{c} \in [0,1]^L, \sum_{k=1}^{L} \vec{c}_k = 1,\\
                    & \vec{t} \in \{0,1\}^{L}.
\end{aligned}
\end{equation}

Due to the rearrangement inequality, we can prove that in the optimal solution,
\begin{equation}
    \max(\alpha,\vec{c}^*_1) \geq \max(\alpha,\vec{c}^*_2) \geq \cdots \geq \max(\alpha,\vec{c}^*_L).
\end{equation}
Then by using Theorem 3, there must exist a $k^* \in \{1,\dots,L\}$ such that
\begin{equation} \label{eq:opt_t}
\begin{aligned}
    &\vec{t}^*_i = 0, i = 1,\dots,k^* \\
    &\vec{t}^*_i = 1, i = k^*+1,\dots,L.
\end{aligned}
\end{equation}
This also implies that only the first $k^*$ elements of $\vec{c}^*$ have nonzero values (greater than $\alpha$ actually). Furthermore, if $k^*$ is known, finding the optimal $\vec{c}^*$ is to solve the following problem:
\begin{equation}
\begin{aligned}
\max_{\vec{c}} \quad & E=\sum_{k=1}^{k^*} m_k \log \mat{c}_k,\\
\text{subject to} \quad & \vec{c} \in [0,1]^L, \sum_{k=1}^{k^*} \vec{c}_k = 1,
\end{aligned}
\end{equation}
whose solution is
\begin{equation} \label{eq:opt_c}
\begin{aligned}
    &\vec{c}^*_i = \frac{m_i}{\sum_{k=1}^{k^*}{m_k}}, i = 1,\dots,k^*,\\
    &\vec{c}^*_i = 0, i = k^*+1,\dots,L.
\end{aligned}
\end{equation}

The above analysis leads to a simple algorithm. We enumerate $k^*$ from 1 to $L$. For each $k^*$, $\vec{t}^*$ and $\vec{c}^*$ are computed by using Eq~\eqref{eq:opt_t} and~\eqref{eq:opt_c}, respectively. Then we evaluate the objective function $E$ and record the best solution.

% subsection confusion_estimation (end)

% section label_noise_model (end)

\section{Deep Learning from Noisy Labels} % (fold)
\label{sec:deep_learning_from_noisy_labels}

\begin{figure*}
\begin{center}
\includegraphics[width=0.9\linewidth]{figure/diagram.pdf}
\end{center}
\caption{System diagram of our method. Two CNNs are used to predict the class label $p(\vec{y}|\vec{x})$ and the noise type $p(\vec{z}|\vec{x})$, respectively. The label noise model layer infers the true label according to the predictions and the noisy label, which is then used to supervise the training of CNNs. Data with clean labels are also mixed in to prevent the models from drifting away.}
\label{fig:framework}
\end{figure*}

We integrate the proposed label noise model into a deep learning framework. As demonstrated in Figure~\ref{fig:framework}, we predict the probability $p(\vec{y}|\vec{x})$ and $p(\vec{z}|\vec{x})$ by using two independent CNNs. Moreover, we append a label noise model layer at the end, which takes as input the CNNs' prediction scores and the observed noisy label. Stochastic Gradient Ascent with backpropagation technique is used to approximately optimize the whole network. In each forward pass, the label noise model layer computes the posterior of latent variables according to Eq~\eqref{eq:posterior}. While in the backward pass, it computes the gradients according to Eq~\eqref{eq:gradient}.

% The forward procedure is summarized in Algorithm~\ref{alg:forward}, and the backward pass for computing gradient with respect to $p(\vec{y}|\vec{x})$ is illustrated in Algorithm~\ref{alg:backward}.

% \begin{algorithm}
%     \SetKwData{Sum}{sum}
%     \SetKwFunction{Softmax}{Softmax}\SetKwFunction{Q}{Q}
%     \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}

%     \Input{Observed weak label vector $\tilde{\vec{y}}$ of size $N$}
%     \Input{Object class scores matrix $Y$ of size $N \times L$}
%     \Input{Noise type scores matrix $Z$ of size $N \times 3$}
%     \Output{Posterior probability matrix $P$ of size $L \times 3$}
%     \BlankLine

%     $P_y \leftarrow$ \Softmax{$Y$}\;
%     $P_z \leftarrow$ \Softmax{$Z$}\;
%     \BlankLine

%     \emph{\Q is the function that implements Eq~\eqref{eq:likelihood}}\;
%     \BlankLine

%     \For {$i \leftarrow 1$ \KwTo $N$} {
%         \Sum $\leftarrow 0$\;
%         \For {$j \leftarrow 1$ \KwTo $L$} {
%             \For {$k \leftarrow 1$ \KwTo $3$} {
%                 $P[i,j,k] \leftarrow P_y[i,j] \times P_z[i,k] \times$ \Q{$\tilde{\vec{y}}$,j,k}\;
%                 \Sum $\leftarrow$ \Sum$+ P[i,j,k]$\;
%             }
%         }
%         $P[i,:,:] \leftarrow P[i,:,:] /$ \Sum\;
%     }

%     \caption{Forward pass}
%     \label{alg:forward}
% \end{algorithm}
% \begin{algorithm}
%     \SetKwData{Sum}{sum}
%     \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}

%     \Input{Object class scores matrix $Y$ of size $N \times L$}
%     \Input{Posterior probability matrix $P$ of size $L \times 3$}
%     \Output{Gradient matrix $\Delta_Y$ of size $N \times L$}
%     \BlankLine

>>>>

% Use class option [extendedabs] to prepare the 1-page extended abstract.
\documentclass[extendedabs]{bmvc2k}
\usepackage[colorlinks = true,
            linkcolor = blue,
            urlcolor  = blue,
            citecolor = blue,
            anchorcolor = blue]{hyperref}

% Document starts here
\begin{document}

\title{Learning From Massive Noisy Labeled Data for Image Classification}
\addauthor{
Tong Xiao$^1$, 
Tian Xia$^2$,
Yi Yang$^2$,
Chang Huang$^2$,
and Xiaogang Wang$^1$
}{}{1}
\addinstitution{
$^1$The Chinese University of Hong Kong.
$^2$Baidu Research.
}

\maketitle
\let\thefootnote\relax\footnote{This is an extended abstract. The full paper is available at the \href{http://www.cv-foundation.org/openaccess/CVPR2015.py}{Computer Vision Foundation webpage}. }
\vspace{-0.2in}

\noindent
Deep learning with large-scale supervised training data has shown very impressive improvement on image classification challenge~\cite{krizhevsky2012imagenet}. However, it requires reliable annotations from millions of images which are often expensive and time-consuming to obtain~\cite{deng2009imagenet}, preventing deep models from being quickly trained on new image recognition problems. Thus it is necessary to develop new efficient labeling and training frameworks for deep learning.

One possible solution is to automatically collect annotations from the Internet web images (\emph{i.e.} extracting tags from the surrounding texts) and directly use them as ground truth to train deep models. Unfortunately, these labels are extremely unreliable due to various types of noise (\emph{e.g.} labeling mistakes from annotators or computing errors from extraction algorithms). Many works have shown that noisy labels could adversely impact the classification accuracy of the induced classifiers~\cite{nettleton2010study}. Various label noise-robust algorithms~\cite{manwani2013noise} and data cleansing algorithms~\cite{brodley2011identifying} are developed but experiments show that performances are still affected by label noise~\cite{bartlett2006convexity}.

Although annotating all the data is costly, it is often easy to obtain a small amount of clean labels. Semi-supervised learning methods~\cite{lee2013pseudo} can also be employed by simply discarding all the noisy labels. Researchers have also studied the transferability of Convolutional Neural Networks (CNNs) by finetuning an ImageNet pretrained model on smaller datasets of specific tasks~\cite{oquab2014learning}. However, these methods cannot fully utilize the massive noisy labeled data, which may lead to suboptimal results.

@@

To cope with this challenge, we extend CNNs with a novel probabilistic model, which infers the true labels and uses them to supervise the training of the network. Our work is inspired by~\cite{sukhbaatar2014learning}, which modified a CNN by inserting a linear layer on top of the softmax layer to map clean labels to noisy labels. However,~\cite{sukhbaatar2014learning} assumed noisy labels are conditionally independent with input images given clean labels. By examining the collected dataset, we find that this assumption is too strong to fit real-world data well. For example, in Figure~\ref{fig:noise_cluster}, all the images should belong to ``Hoodie''. The top five are correct while the bottom five are either mislabeled as ``Windbreaker'' or ``Jacket''. Since different sellers have their own bias on different categories, they may provide wrong keywords for similar clothes. We can capture these visual patterns to estimate how likely an image is mislabeled. Based on these observations, we further introduce two types of label noise:
\begin{itemize}
\item \textbf{Confusing noise} makes the noisy label reasonably wrong. It usually occurs when the image content is confusing (\emph{e.g}, the samples with ``?'' in Figure~\ref{fig:overview}).
\item \textbf{Pure random noise} makes the noisy label totally wrong. It is often caused by either the mismatch between an image and its surrounding text, or the false conversion from the text to label (\emph{e.g.}, the samples with ``$\times$'' in Figure~\ref{fig:overview}).
\end{itemize}

Our proposed probabilistic model builds the relations among images, noisy labels, ground truth labels, and noise types, where the latter two are treated as latent variables. We use the Expectation-Maximization (EM) algorithm to solve the problem and integrate it into the training process of CNNs. Experiments on the collected dataset show that our model can better detect and correct the wrong labels, which benefits the training of underlying CNNs.

\begin{figure}[t]
\begin{center}
\includegraphics[width=0.8\linewidth]{figure/overview.pdf}
\end{center}
\caption{Overview of our approach. Labels of web images often suffer from different types of noise. A label noise model is proposed to detect and correct the wrong labels. The corrected labels are used to train underlying CNNs.}
\label{fig:overview}
\end{figure}

\begin{figure}[t]
\begin{center}
\includegraphics[width=0.8\linewidth]{figure/noise_cluster.pdf}
\end{center}
\caption{Mislabeled images often share similar visual patterns.}
\label{fig:noise_cluster}
\end{figure}

\bibliography{extbib}
